---
title: "Reproducible Project 1"
author: "K Cousar"
date: "8/3/2020"
output: html_document
---

This document contains my work and answers to Coursera Reproducible Research Project 1. Data download and preperation is shown below. 

```{r}
data <- read.csv("activity.csv")
head(data)
str(data)
summary(data)
```

## Question 1  
1. Calculate the total number of steps per day.
```{r}
steps_perday <- aggregate(steps~date, data, sum, na.rm=TRUE)
```
2. Make a historgram of total number of steps per day. 
```{r}
hist(steps_perday$steps)
```  
3. Report the mean and median of total number of steps per day
```{r}
summary(steps_perday$steps)
```

## Question 2  
1. Make a time series plot of the 5-minute intervals and the average number of steps taken, averaged across all days.
```{r}
steps_perinterval <- aggregate(steps~interval, data=data, mean, na.rm=TRUE)
plot(steps~interval, data=steps_perinterval, type = "l")
```    
2. Report the interval with the maximum number of steps taken per day.
```{r}
interval_maxsteps <- steps_perinterval[which.max(steps_perinterval$steps),]$interval
interval_maxsteps
```

## Question 3  
1. Report the number of missing values.
```{r}
missing <- is.na(data)
sum(missing)
```  
2. Create a new dataset without missing values. I used the means for each interval. 
```{r}
means_forNA <- function(interval){
  steps_perinterval[steps_perinterval$interval==interval,]$steps
}
data2 <- data
  for (i in 1:nrow(data2)) {
    if(is.na(data2[i,]$steps)){
      data2[i,]$steps <- means_forNA(data2[i,]$interval)
  }
}
```  
3. Make a histogram of the total number of steps taken each day.
```{r}
steps_perday2 <- aggregate(steps~date, data=data2, sum)
hist(steps_perday2$steps)
```  
4. Report the mean and median of the new dataset for number of steps taken per day
```{r}
summary(steps_perday2$steps)
```

## Question 4  
1. Create a new factor variable to indicate if the day is a weekend of a weekday.
```{r}
data2$date <- as.Date(strptime(data2$date, format = "%Y-%m-%d"))
data2$day <- weekdays(data2$date)
  for (i in 1:nrow(data2)) {
    if(data2[i,]$day %in% c("Saturday", "Sunday")) {
      data2[i,]$day <- "weekend"
    }
    else{
      data2[i,]$day <- "weekday"
    }
  }
```  
2. Create a panel plot showing the time series of 5-minute intervals and average number of steps taken, respectfully for both weekends and weekdays. 
```{r}
steps_byday <-aggregate(data2$steps~data2$interval+data2$day, data = data2, mean)
names(steps_byday) <- c("Interval", "Day", "Steps")
library(lattice)
xyplot(Steps~Interval|Day, steps_byday, type = "l", layout=c(1,2), xlab= "Interval", ylab = "Number of Steps")
```  
# This concludes this assignment. 